Prosper Loan Data Analysis by Preston Hall
========================================================
  

> We will be using the data provided by Prosper Bank loans. Prosper Bank has
facilitated more than $12 billion in loans to more than 770,000 people. You can
find more information about Prosper Bank at their [website.](www.prosper.com)


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(stringr)
library(plyr)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data
loanData <- read.csv('prosperLoanData.csv')
```

> Introduction: (Needs work) I have worked in the financial industry for close to
10 years and have serviced and reviewed many loan documents. I am excited to 
analyze this data and discover new information.

> The data that we are exploring is provided in a .csv file 86.5 MB. There are a
total of 81 variables and 113,937 observations. There is information than what is 
needed for this project so I will subest it out to 16 different varaiables and 
remove all of the NA values. 

```{r echo=FALSE, message=FALSE, warning=FALSE, SubsetTheDate}
loanSubset <- subset(loanData, select = c(
  Term, 
  LoanStatus,
  BorrowerRate,
  BorrowerState,
  Occupation,
  EmploymentStatus,
  CreditScoreRangeLower,
  CreditScoreRangeUpper,
  CreditGrade,
  OpenCreditLines,
  CurrentDelinquencies,
  AmountDelinquent,
  DebtToIncomeRatio,
  StatedMonthlyIncome,
  IncomeRange,
  MonthlyLoanPayment
))
```

> Now that I have the subset that I want to work with, I will clean it up a bit by removing a few of the NA values and adding factors to IncomeRange, Term and CreditGrade. IncomeRange and CreditGrade will ibe ordered factors.

```{r echo=FALSE, message=FALSE, warning=FALSE, CleanAndRemoveNA}
loanSubset$IncomeRange <- factor(loanSubset$IncomeRange, 
                                 levels=c("$1-24,999", "$25,000-49,999", 
                                          "$50,000-74,999", "$75,000-99,999", 
                                          "$100,000+","Not employed"), 
                                 ordered=TRUE, 
                                 exclude = "Not displayed")

loanSubset$Term <- as.factor(loanSubset$Term)

loans <- na.exclude(loanSubset)

loans$CreditGrade <- factor(loans$CreditGrade, 
                            levels = c("AA", "A", "B", "C", "D", "E", "HR", "NC"), 
                            ordered = TRUE)
```

# Univariate Plots Section

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
  There should be multiple code chunks and text sections; the first one below is
just to help you get started.


### Variables I will be working with
```{r Variable Names, echo=FALSE, message=TRUE, warning=TRUE}
names(loans)
```

- Term
- LoanStatus
- BorrowerRate
- BorrowerState
- Occupation
- EmploymentStatus
- CreditScoreRangeLower
- CreditScoreRangeUpper
- CreditGrade
- OpenCreditLines
- CurrentDelinquencies
- AmountDelinquent
- DebtToIncomeRatio
- StatedMonthlyIncome
- IncomeRange
- MonthlyLoanPayment

### Borrower APR
```{r echo=FALSE, message=FALSE, warning=FALSE, BorrowerAPR}
qplot(x = BorrowerRate, data = loans, binwidth = 0.005) +
  scale_x_continuous(limits = c(0,0.36), breaks = seq(0,0.36,0.02)) +
  scale_y_continuous(breaks = seq(0,6000,500))
summary(loans$BorrowerRate)
```

> This graph is interesting as it is has a fairly normal distribution save for the rate at 32% the amount jumps up to over 4500.


### Stated Monthly Income
```{r echo=FALSE, message=FALSE, warning=FALSE, StatedMonthlyIncome}
summary(loans$StatedMonthlyIncome)
income1 <- ggplot(aes(x = StatedMonthlyIncome), data = loans) +
  geom_histogram(binwidth = 500) +
  xlim(0, 20000) +
  xlab('Stated Monthly Income in USD') +
  ylab('Number of Applicants')

income2 <- ggplot(loans, aes(x = StatedMonthlyIncome)) +
  geom_density() +
  xlim(0, 20000)

grid.arrange(income1, income2, ncol = 1)
```
> You can see on this graph that the plot peaks at about $3,000 per month and then decreases. I had to limit the x axis as there is an outlier that skews the graph to a point that is unreadable. 


### Term

```{r echo=FALSE, message=FALSE, warning=FALSE, Term}
summary(loans$Term)

qplot(Term, data = loans)
```

> The types of loans fall into 3 different terms 12, 36, or 60 months. Or 1, 3, or 5 years respectively. The most common term is 3 years with over 73,000 loans.

### Credit Score effect on Rate

```{r echo=FALSE, message=FALSE, warning=FALSE, CreditVRate}
CreditRate1 <- ggplot(aes(x = CreditScoreRangeUpper, y = BorrowerRate), data = loans) +
  geom_point()

CreditRate2 <- ggplot(aes(x = CreditScoreRangeUpper, y = BorrowerRate), data = loans) +
  geom_jitter(alpha = 1/50, position = position_jitter(h = 0)) 

CreditRate3 <- ggplot(aes(x = CreditScoreRangeUpper, y = BorrowerRate), data = loans) +
  geom_jitter(alpha = 1/50, position = position_jitter(h = 0)) +
  geom_smooth()

grid.arrange(CreditRate1, CreditRate2, CreditRate3)
```

> This scatter plot was very difficult to see any relation between Credit Score and Borrower Rate. I transformed the data so the plotted points were more transparent and added a linear regression line. You can see that the the interest rate decreases as the credit score improves. 

### CreditGrade

```{r echo=FALSE, message=FALSE, warning=FALSE, CreditGrade}
qplot(data = loans, x = CreditGrade) +
  xlim("AA", "A", "B", "C", "D", "E", "HR")
summary(loans$CreditGrade)
```

> Customer are given a credit grade when provided a loan by Prosper Bank. AA would be the highest score with HR (High Risk) being at the lower end. Most customs have a score between B and D. I am not including NA values in this graph.


# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?
```{r}
dim(loanData)
# structure(loanData)
```

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?
  
  ### Did you observe any interesting relationships between the other features \
  (not the main feature(s) of interest)?
  
  ### What was the strongest relationship you found?
  
  
  # Multivariate Plots Section
  
  > **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
                                                                         units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------
  
  # Reflection
  
  > **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!