Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:
```{r}
library(ggplot2)
library(dplyr)
library(gridExtra)
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
```

***

### Scatterplots
Notes:

```{r Scatterplots}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point()
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point() + 
  xlim(13, 90) 
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha = 1/20) + 
  xlim(13, 90) 
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +  #go back to geom_point and add the position arguement to prevent negative numbers
  xlim(13, 90) +
  coord_trans(y = 'sqrt')
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Notes:
Examine the relationship between friendships_initiated (y) and age (x) using the ggplot syntax.

We recommend creating a basic scatter plot first to see what the distribution looks like. and then adjusting it by adding one layer at a time.

What are your observations about your final plot?

Remember to make adjustments to the breaks of the x-axis and to use apply alpha and jitter.

** Remember ** When using coord_trans, some values may be 0 so make sure to update to geom_point to geom_jitter with position = position_jitter with a height of 0 to account for negative values. 

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_jitter(alpha = 1/10, position = position_jitter(h = 0)) +
  coord_trans(y = 'sqrt')
```

***

### Overplotting and Domain Knowledge
Notes: Video 8. Overplotting and Domain Knowledge: 
Used percentages to show a better perspective based on the amount of friends each surveyor had rather than the friend count. Gives a better relational answer to the question.

***

### Conditional Means
Notes: Using package dplyr

dplyr allows us to split up the data frame and apply a function on some parts of the data. Some common functions are:
filter()
group_by()
mutate()
arrange() 

```{r Conditional Means}
# library(dplyr)

age_groups <- group_by(pf, age) # grouped_by age from pf dataset
pf.fc_by_age <- summarise(age_groups,   # creating mean, median and count
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n()) # can only be used for summarize and returns
                                   # number of users
pf.fc_by_age <- arrange(pf.fc_by_age, age) # arranges by age

head(pf.fc_by_age)
```

Create your plot!
** Creates the same plot as above but using the %>% chain (%.% is deprecated)
```
pf.fc_by_age2 <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>% # can only be used for summarize and returns
                     # number of users
  arrange(age)

head(pf.fc_by_age2, 20)
```
```{r Conditional Means Plot}
# **My code**
# qplot(data = pf.fc_by_age, x = age, y = friend_count_mean)    

# Video code
ggplot(aes(age, friend_count_mean), data = pf.fc_by_age) + 
  geom_line() #geom_point() will give the scatterplot similar to above
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) +
  coord_cartesian(xlim = c(13, 70), ylim = c(0, 1000), expand = "FALSE") + # allows me to zoom in on the graph
  geom_point(alpha = 0.05,
             position = position_jitter(h = 0),
             color = 'orange') +
  coord_trans(y = 'sqrt') +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = quantile, 
            fun.args = list(probs = .1), 
            linetype = "dashed", color = "blue") +
  geom_line(stat = 'summary', fun.y = quantile, 
            fun.args = list(probs = .5), 
            color = "blue") +
  geom_line(stat = 'summary', fun.y = quantile, 
            fun.args = list(probs = .9), 
            linetype = "dashed", color = "blue")
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes: Try to summarize the strength of the correlation in a single line. ?core.test 

A correlation greater than 0.3 or less than -0.3 is meaningful but small, 0.5 is moderate and 0.7 is large.

```{r Correlation}
cor.test(pf$age, pf$friend_count,
         method = "pearson")
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age < 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
  geom_point()
```

***

### Strong Correlations
Notes: Limit the x and y limit from above to show a stronger correlation

```{r Strong Correlations}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95))
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(pf$www_likes_received, pf$likes_received, method = "pearson")

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
# install.packages('alr3')
library(alr3)
data(Mitchell)
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot. 0

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Temp, Mitchell$Month, method = "pearson")
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point() + 
  scale_x_continuous(breaks = seq(0, 203, 12))
```

***

### A New Perspective

What do you notice?
Response: If you stretch the plot you can see the change in the weather

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
pf$age_with_months <- with(pf, pf$age + (1 - pf$dob_month / 12))
```

***

### Age with Months Means

```{r Age with Months Means}
pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age_with_months)

head(pf.fc_by_age_months)
```

Programming Assignment - Subset to only include ages under 71
```{r Programming Assignment}
ggplot(aes(x = age_with_months, y = friend_count_mean), 
       data = subset(pf.fc_by_age_months, age_with_months < 71)) + 
  geom_line() 
```

***

### Noise in Conditional Means
Example of the bias variance tradeoff

```{r Noise in Conditional Means}
p1 <- ggplot(aes(x = age, y = friend_count_mean), 
       data = subset(pf.fc_by_age, age < 71)) + 
  geom_line() +
  geom_smooth()

p2 <- ggplot(aes(x = age_with_months, y = friend_count_mean), 
       data = subset(pf.fc_by_age_months, age_with_months < 71)) + 
  geom_line()  +
  geom_smooth()

p3 <- ggplot(aes(x = round(age/5) * 5, y = friend_count),
             data = subset(pf, age < 71)) +
  geom_line(stat = 'summary', fun.y = mean)

grid.arrange(p2, p1, p3, ncol = 1)
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

