Lesson 3
========================================================

***

# Reading in the date
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
pf <- read.csv("pseudo_facebook.tsv", sep = "\t")
names(pf)
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
# install.packages('ggplot2')
library(ggplot2)
# install.packages('ggthemes', dependencies = TRUE)
library(ggthemes)
```


```{r}
qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks=1:31)

```

***
```{r}
ggplot(aes(x = dob_day), data = pf) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31)
```

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:
```{r}
ggplot(data = pf, aes(x = dob_day)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month) #create histogram for each category
```

***

### Estimating Your Audience Size
Notes: This creates a histogram of the friend counts
***
```{r}
qplot(x = friend_count, data = pf)

```
# You can limit the amount of axes with this code
```{r}
qplot(x = friend_count, data = pf, xlim = c(0, 1000))
```
# OR
```{r}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0,1000, 50)) +
  facet_wrap(gender~.)

```
# check how many males vs females there are
```{r}
table(pf$gender) # To get a count(add below to get average)
by(pf$friend_count, pf$gender, summary)
```
# The parameter color determines the color outline of objects in a plot.
# The parameter fill determines the color of the area inside objects in a plot.
# You might notice how the color black and the hex code color of #099DD9 (a shade of blue) are wrapped inside of I(). The I() functions stand for 'as is' and tells qplot to use them as colors.
```{r}
ggplot(aes(x = tenure), data = pf) +
  geom_histogram(binwidth = 30, color = 'black', fill = '#099DD9')

ggplot(aes(x = tenure / 365), data = pf) +
  geom_histogram(color = 'black', fill = '#F79420') +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7)) +
  xlab('Number of years using Facebook') +
  ylab('Number of users in sample')
```

```{r}
ggplot(aes(x = age), data = pf) +
  geom_histogram(binwidth = 1, fill = '#5760AB') +
  scale_x_continuous(breaks = seq(0, 113, 5))
```


#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}

```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}

```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}

```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}

```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}

```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

```{r Tenure}

```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}

```

***

### Labeling Plots
Notes:

```{r Labeling Plots}

```

***

### User Ages
Notes:

```{r User Ages}

```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:
```{r}
qplot(x = friend_count, data = pf)

summary(pf$friend_count)

summary(log10(pf$friend_count + 1))

summary(sqrt(pf$friend_count))
```
```{r}
# install.packages('gridExtra')
library(gridExtra)
```
```{r}
# define individual plots
p1 = qplot(x = friend_count, data = pf) 
p2 = qplot(log10(pf$friend_count + 1)) 
p3 = qplot(sqrt(pf$friend_count))  
# arrange plots in grid
grid.arrange(p1, p2, p3, ncol=1)
```
# same as above but different syntax
```{r}
p4 <- ggplot(aes(x = friend_count), data = pf) + geom_histogram()
p5 <- p4 + scale_x_log10()
p6 <- p4 + scale_x_sqrt()

grid.arrange(p4, p5, p6, ncol=1)
```


***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}

```

***


### Frequency Polygons

```{r Frequency Polygons}
qplot(x = friend_count, y = ..density../sum(..density..), 
      data = subset(pf, !is.na(gender)), 
      binwidth = 10, 
      geom = 'freqpoly', 
      color = gender) +
  scale_x_continuous(limits = c(0, 1000), 
                     breaks = seq(0, 1000, 50)) +
  xlab('Friend Count') +
  ylab('Proportion of users with that friend count')
```

***
# Similar ggplot syntax
# |
# ggplot(aes(x = friend_count, y = ..count../sum(..count..)),
#       data = subset(pf, !is.na(gender))) +
#  geom_freqpoly(aes(color = gender), binwidth=10) +
#  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
#  xlab('Friend Count') +
#  ylab('Proportion of users with that friend count')

# Equivalent ggplot syntax for solution video:
# |
# ggplot(aes(x = www_likes), data = subset(pf, !is.na(gender))) +
#  geom_freqpoly(aes(color = gender)) +
#  scale_x_log10()
### Likes on the Web
Notes:

```{r Likes on the Web}
 ggplot(aes(x = www_likes), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) +
  scale_x_log10() +
  xlab('Likes') +
  ylab('Proportion of likes per gender')
```

```{r}
by(pf$www_likes, pf$gender, sum)
```

***

### Box Plots
Notes:

```{r Box Plots}
qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot') +
      coord_cartesian(ylim = c(0, 250)) 

by(pf$friend_count, pf$gender, summary)
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
summary(pf$mobile_likes)

summary(pf$mobile_likes > 0)

mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)
```

Response:
```{r}
sum(pf$mobile_check_in == 1)/length(pf$mobile_check_in)
```

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!